---
title: "DATA 606 Data Project Proposal"
author: "Yohannes Deboch"
---

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data
cars <- read.table("https://archive.ics.uci.edu/ml/machine-learning-databases/auto-mpg/auto-mpg.data", header = FALSE)
colnames(cars) <- c("mpg","cylinders","displacement","horsepower","weight","acceleration","model year","origin","car name")
```


### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

Answer: 

This Project is aimed at determining cars millage based on other attribures . The central question that will  be answered by this project is : Can we predict cars mileage based on other attributes?


### Cases 

In this project the cases are  observation of a cars information.

```{r}
nrow(cars)
```
There were 397 cases. 


### Data collection 

The data was collected from the UCI machine learning repository website.
Data set name: 
Auto MPG Data Set 


### Type of study 

 This is a observational study.


### Data Source 

Data link 
https://archive.ics.uci.edu/ml/datasets/Auto+MPG


### Dependent Variable

Mile per gallon (mpg) is the dependent variable. It's continuous.


### Independent Variable

 The independent variables are the following ones
```{r}
colnames(cars)[2:8]
```


### Relevant summary statistics

First and last few rows of the dat
```{r}
head(cars)
tail(cars)
```

Structure of dataset
```{r}
str(cars)
```



```{r}
summary(cars)
```

```{r}
boxplot(cars$mpg~cars$cylinders,main = "Milage by Cylinders")
```

```{r}
boxplot(cars$mpg~cars$origin,main = "Milage by origin")
```

```{r}
boxplot(cars$mpg~cars$cylinders, main = "Milage by cylinders")
```
```{r}
plot(cars$mpg,cars$weight,main = "Milage by weight")
```
```{r}
hist(cars$mpg,main="Distribution of Milage")
```
```{r}
cor(cars[,c(1,2,3,5,6,7)])
```

From the summary statistics we can see that there are no missing values in the data. mpg response variable is normally distributed . From correlation we can see that mpg is strongly correlated with cylinders,displacement,weight,acceleration and model year are  useful in predicting mpg.